MS_Imp_Guidelines

### 2023-01-31 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
edits up to mid discussion

(i) the ready4 framework, that is comprised of:
  - a set of standards for implementing an accountable, reusable and updatable MOSHEM; and
  - a modelling toolkit (online repositories and software) for applying those standards to the authoring of MOSHEM modules, datasets and analyses; and

(ii) how we used the ready4 framework to develop and apply modules for utility mapping.

>>>>

(i) a framework that specifies MOSHEM standards and provides the tools to implement them; and

(ii) the application of that framework to develop and use an initial set of ready4 modules in a utility mapping study.

### 2023-01-31 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd
edits up to mid discussion

In this article we described a framework that we developed to help us implement ready4 - a MOSHEM in youth mental health. We outlined framework standards for an accountable, reusable and updatable MOSHEM and described the modelling toolkit we created for applying those standards to the development and use of the ready4 MOSHEM. We also provided an overview of an initial set of MOSHEM modules developed with the framework to implement a utility mapping study. We reviewed the modules, datasets and analyses generated by that study against framework standards. The work we have described has potential implications for the development of the ready4 MOSHEM and for health economic modelling in mental health and highlights more general issues for health economic modellers and funders of health economic research.

## Implications for implementing ready4
The most direct implication of the development of the ready4 framework is that it makes it feasible for us to implement a MOSHEM in youth mental health. The standards specified by the framework have enabled us to partially automate workflows for developing and applying ready4 through use of the framework's modelling toolkit. We have demonstrated the practical utility of the modelling toolkit by applying it to authoring, documenting and disseminating ready4 module libraries  [@hamilton_matthew_2022_6084467; @hamilton_matthew_2022_6084824; @hamilton_matthew_2022_6116701; @gao_caroline_2022_6130155], datasets [@DVN/HJXYKQ_2021; @DVN/DKDIB0_2021]; analyses [@hamilton_matthew_2022_6129906; @hamilton_matthew_2022_6416330; @hamilton_matthew_p_2022_6321821], reporting templates [@hamilton_matthew_2022_6116385; @matthew_p_hamilton_2022_5976988] and prediction tools  @matthew_p_hamilton_2021_5646669] used in a utility mapping study [@Hamilton2021.07.07.21260129]. The standardised and partially automated workflows used in creating and sharing these artefacts has the potential to generate significant efficiencies as we apply the ready4 framework to undertaking new studies.

We have also been able to demonstrate the interoperability of the initial ready4 modules developed with the modelling toolkit. The program used to implement the utility mapping analysis [@hamilton_matthew_2022_6129906] combines modules from four module libraries ([@hamilton_matthew_2022_6084467; @hamilton_matthew_2022_6084824; @hamilton_matthew_2022_6116701; @gao_caroline_2022_6130155]) and two framework libraries [@ready4use2022; @ready4show2022]. Example literate programs published on the ready4 documentation website [@rfwn2022] use toy data [@DVN/HJXYKQ_2021] to illustrate the potential for ready4 modules to facilitate study replication and transferability. The framework's standards also provide a mechanism for to assess the extent to which the ready4 MOSHEM meets explicit accountability, reusability and updatability goals.

@@

Currently all the framework and module libraries we have developed are available only as "development" releases. An early priority for us is to undertake the additional development, testing and documenting of these libraries so that we can submit production versions of each library to CRAN [@CRAN2022]. Making these libraries available on CRAN is a likely prerequisite of high levels of reuse. Similarly, completing our current work program will create new tools for modelling help-seeking choice, spatial epidemiology, household populations and primary mental health services that may enhance the potential usefulness of the ready4 MOSHEM to health economists working in youth mental health.

## Implications for economic modelling in mental health
Open source approaches have been recommended to help develop the mental health modelling field [@RN73] but only one mental health related model (in Alcohol Use Disorder [@Basu2018]) is currently indexed in the Open Source Models Clearinghouse [@OSMC_20xx;@Emerson2019]. We are aware of just one other open source mental health model - a reference model in Major Depressive Disorder - that is currently in development [@IVIMDD2022]. Of the known barriers to adoption of open source models by health economists (including issues like intellectual property and confidentiality [@Pouwels2022; @Wu2019]), our experience suggests that the enormous effort that is required to prepare model code and data for public release in formats that facilitate appropriate reuse by third parties and to maintain and improve these artefacts may be the biggest challenge.

Automated tools such as those we developed in our modelling toolkit can help reduce the burden of some of these tasks. However, we think the current low rates of adoption by health economists of open source approaches [@Emerson2019; @Michalczyk2018; @Feenstra2022] (which in turn facilitate the collaboration that make modular models more attractive) is unlikely to change soon without significant and strategic changes by research funders. The incentive structures that health economists work within do not promote the investment of significantly more time on each study to enable peers to reuse their work.

Reducing waste in research is a responsibility of research funders [@chalmers2014increase] and the poor reproducibility [@Jalali2021; @McManus2019; @Bermejo2017], limited reusability [@Emerson2019; @Michalczyk2018; @Feenstra2022] and uncertain validity [@Ghabri2019; @Radeva2020] of health economic models is wasteful. Approximately 4,000 mental health focused economic evaluation reports were produced between 2000 to 2019 [@RN34]. The intellectual asset represented by this literature could be very substantial if many of the models described in these reports could be brought (and kept) up to date and made available in formats that maximised transferability to diverse decision contexts. We believe that modular and open source approaches would be well suited to accomplishing this goal and that the framework we have developed could act as an early prototype for solving some of the technical challenges of this task. Ideally such a programme of research would be resourced to be sustained over the medium to long term and to engage a diverse network of investigators, contributors and advisers from high, middle and low income countries.

@@

@@

<!-- (described in more detail on the project documentation website [@rfwn2022])  -->

@@

>>>>

In this article we described a framework that we developed to help us implement ready4 - a MOSHEM in youth mental health. We outlined framework standards for an accountable, reusable and updatable MOSHEM and described the modelling toolkit we created for applying those standards to the development and use of the ready4 MOSHEM. We also provided an overview of an initial set of MOSHEM modules developed with the framework to implement a utility mapping study. We reviewed the modules, datasets and analyses generated by that study against framework standards. The work we have described has potential implications for the development of the ready4 MOSHEM and for health economic modelling in mental health. A number of issues have more general relevance to health economic modellers and funders of health economic research.

## Implications for implementing ready4
The most direct implication of the development of the ready4 framework is that it makes it feasible for us to implement a MOSHEM in youth mental health. The standards specified by the framework have enabled us to partially automate workflows for developing and applying ready4 through use of the framework's modelling toolkit. We have demonstrated the practical utility of the modelling toolkit by applying it to authoring, documenting and disseminating ready4 module libraries  [@hamilton_matthew_2022_6084467; @hamilton_matthew_2022_6084824; @hamilton_matthew_2022_6116701; @gao_caroline_2022_6130155], datasets [@DVN/HJXYKQ_2021; @DVN/DKDIB0_2021]; analyses [@hamilton_matthew_2022_6129906; @hamilton_matthew_2022_6416330; @hamilton_matthew_p_2022_6321821], reporting templates [@hamilton_matthew_2022_6116385; @matthew_p_hamilton_2022_5976988] and prediction tools  @matthew_p_hamilton_2021_5646669] used in a utility mapping study [@Hamilton2021.07.07.21260129]. The standardised and partially automated workflows used in creating and sharing these artefacts has the potential to generate significant efficiencies as we apply the ready4 framework to undertaking new economic studies.

We have also been able to demonstrate the interoperability of the initial ready4 modules developed with the modelling toolkit. The program used to implement the utility mapping analysis [@hamilton_matthew_2022_6129906] combines modules from four module libraries ([@hamilton_matthew_2022_6084467; @hamilton_matthew_2022_6084824; @hamilton_matthew_2022_6116701; @gao_caroline_2022_6130155]) and two framework libraries [@ready4use2022; @ready4show2022]. Example literate programs published on the ready4 documentation website [@rfwn2022] use toy data [@DVN/HJXYKQ_2021] to illustrate the potential for ready4 modules to facilitate study replication and transferability. As demonstrated by the checklist we developed (Table \@ref(tab:checktb)), our framework's standards also provide a mechanism for to assess the extent to which the ready4 MOSHEM meets explicit objectives.

@@

To progress from a technical capability to behavioural outcomes, both our framework and MOSHEM need further work. Currently all the framework and model module libraries we have developed are available only as "development" releases. An early priority for us is to undertake the additional development, testing and documenting of these libraries so that we can submit production versions of each library to CRAN [@CRAN2022]. Making R libraries available on CRAN is normally a prerequisite for high levels of reuse.

The transferability claims we make for our existing modules are to date supported only by example programs using toy data. Our future work aims to address this with real world studies that apply modules to different concepts and contexts. Our current work program also aims to create new ready4 modules for modelling help-seeking choice, spatial epidemiology, household populations and primary mental health services that we hope will provide others with more reasons to use ready4 and contribute to its development. To facilitate code contributions by third parties, our libraries for authoring modules [@ready4pack2022: @ready4fun2022; @ready4use2022] require some additional development to make them easier to use by third parties without knowledge of the naming and directory strucuture conventions we use in authoring code.

<!-- Real applications would be better given importance. -->

## Implications for economic modelling in mental health
Open source approaches have been recommended to help develop the mental health modelling field [@RN73] but only one mental health related model (in Alcohol Use Disorder [@Basu2018]) is currently indexed in the Open Source Models Clearinghouse [@OSMC_20xx;@Emerson2019]. We are aware of just one other open source mental health model - a reference model in Major Depressive Disorder - that is currently in development [@IVIMDD2022]. Of the known barriers to adoption of open source models by health economists (including issues like intellectual property and confidentiality [@Pouwels2022; @Wu2019]), our experience suggests that the biggest challenges may be the enormous effort required to first prepare model code and data for public release in formats that facilitate appropriate reuse by third parties and to then maintain and contunually improve potentially large numbers of digital artefacts.

Automated tools such as those we developed in our modelling toolkit can help reduce the burden associated with some of these tasks. However, we think the current low rates of adoption by health economists of open source approaches [@Emerson2019; @Michalczyk2018; @Feenstra2022] (which in turn facilitate the collaboration that make modular models more attractive) will only change slowly unless there is significant and strategic investments made by research funders. Currently, the incentive structures for health economists do not promote dedicating large quantities of time to enabling peers to reuse their work.

Reducing waste in research is a responsibility of research funders [@chalmers2014increase] and the poor reproducibility [@Jalali2021; @McManus2019; @Bermejo2017], limited reusability [@Emerson2019; @Michalczyk2018; @Feenstra2022] and uncertain validity [@Ghabri2019; @Radeva2020] of health economic models is wasteful. Approximately 4,000 mental health focused economic evaluation reports were produced between 2000 to 2019 [@RN34]. The intellectual asset represented by this literature could be very substantial if many of the models described in these reports could be brought and kept up to date and made available in formats that maximised transferability to diverse decision contexts. We believe that modular and open source approaches would be well suited to accomplishing this goal and that the framework we have developed could act as an early prototype for solving some of the technical challenges of this task. Ideally such a programme of research would be resourced to be sustained over the medium to long term and to engage a diverse network of investigators, contributors and advisers from high, middle and low income countries.

@@

<!-- Resilience of open source project needs multiple maintainers -->

@@

<!-- (described in more detail on the project documentation website [@rfwn2022])  -->
<!-- Sustained funding -->

@@

<!-- Diversify skills - move to tools means software engineering, -->
<!-- More collaboration -->
<!-- Role for SIGs -->
<!-- Make grant proposals that give funders the opportunity to fund this work. -->
<!-- Share some code and data at small scale. -->

### 2023-01-31 Markdown/Manuscript/Child_RMDs/MS_Imp_Example.Rmd
edits up to mid discussion

We applied the ready4 framework to develop an initial set of MOSHEM modules, supplying those modules with data and implementing analyses. These tools were created as part of a previously described study [@Hamilton2021.07.07.21260129] to develop utility mapping models appropriate for use in samples of young people presenting to primary mental health services. The ready4 framework's modelling toolkit created the following artefacts:

@@

We assess that these study outputs either wholly or mostly meet 18 out of 20 framework standards(see Table \@ref(tab:checktb)). The two standards where the study outputs currently fall short are in reporting code coverage and including a user-interface. Both these items are scheduled to be addressed when we release production versions of the code libraries.

>>>>

We applied the framework to develop an initial set of ready4 modules, supply those modules with data and implement modelling analyses. These outputs were created as part of a previously described study [@Hamilton2021.07.07.21260129] to develop utility mapping models appropriate for use in samples of young people presenting to primary mental health services. The ready4 framework's modelling toolkit created the following artefacts:

@@

We created a checklist (Table \@ref(tab:checktb)) that we used to assess the extent to which study outputs met framework standards. We assessed these outputs as wholly or mostly meeting 18 out of 20 standards. The two standards where the study outputs currently fall short are in reporting code coverage and including a user-interface. Both these items are scheduled to be addressed when we release production versions of the code libraries.

### 2023-01-31 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd
edits up to mid discussion

### Standards for an accountable MOSHEM
Guidance on transparency in health economic modelling published over ten years ago [@Eddy2012] made recommendations on documenting models but notably did not include recommendations on sharing model code and data. However, more recent and multidisciplinary healthcare modelling guidance [@Erdemir2020] recommends using existing digital repository services to these types of digital model artefacts. Some types of repository such as GitHub [@github2007] provide tools for disseminating work in progress code and providing highly transparent records of the complete development history and individual authorship contributions of a software project, while others such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection.

@@

Statistical models are a common output of health economic evaluations, but they are often not reported in a format that enables others to confidently and reliably re-use them for out of sample prediction [@Kearns2013]. Open source approaches can help address this by disseminating code artefacts that enable easy and appropriate use of a statistical model to make predictions with new data. However, great care must be exercised when doing so if models are derived from data on human subjects as some software artefacts by default contain a copy of the source dataset. Such dataset copies must therefore be replaced (for example, with synthetic data) and the amended artefact's predictive performance then retested before any public release. Another way to make MOSHEMs easier to use is to develop simple user-interfaces for non-technical users.

### Standards for an updatable MOSHEM
To avoid MOSHEMs going stale - losing validity and usefulness with time - they should be routinely updated.  Each update of code and data should be uniquely identifiable and retrievable, a goal that can be facilitated by use of version control tools [@Erdemir2020] such as [@git20XX].

Potential users of model code and data should be able to easily identify which version is most appropriate to their needs. Using semantic versioning [@semver20xx] conventions can signal the potential importance of an update to users of model code and data. Also informative is to clearly label the type of code hosted in a repository as either:

- "development" (typically the most comprehensive and up to date source code, but potentially not yet sufficiently well documented and tested for potential users to apply to purposes other than testing);

- "production" (code that has been released with a view to it being applied to its stated purposes and therefore typically required to meet defined documentation and testing standards); or

- "archive" (permanent copies of code at key milestones in its development - potentially useful if seeking to use older versions of code to reproduce analyses).

Continuous integration [@CI2017] tools can help verify that each code update passes multiple quality tests. Finally, using deprecation conventions that take an informative and staged approach to retiring old code and data reduces the risk that model revisions have unintended consequences on third party users.

>>>>

### Standards for an accountable MOSHEM
Guidance on transparency in health economic modelling published over ten years ago [@Eddy2012] made recommendations on documenting models but notably did not include recommendations on sharing model code and data. However, more recent and multidisciplinary healthcare modelling guidance [@Erdemir2020] recommends using existing digital repository services to make these types of digital model artefacts publicly available. Some repositories such as GitHub [@github2007] provide tools for disseminating work in progress code and providing highly transparent records of the complete development history and individual authorship contributions of a software project. Other repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection.

@@

Statistical models are a common output of health economic evaluations, but they are often not reported in a format that enables others to confidently and reliably re-use them for out of sample prediction [@Kearns2013]. Open source approaches can help address this by disseminating code artefacts that enable easy and appropriate use of a statistical model to make predictions with new data. However, great care must be exercised when publicly releasing model aretefacts  derived from data on human subjects as they may by default embed a copy of the source dataset. Sensitive dataset copies must therefore be replaced (for example, with synthetic data) and the amended artefact's predictive performance then retested before any public release. Another way to make MOSHEMs easier to use is to develop simple user-interfaces for non-technical users.

### Standards for an updatable MOSHEM
To avoid MOSHEMs going stale - losing validity and usefulness with time - they should be routinely updated.  Each update of code and data should be uniquely identifiable and retrievable, a goal that can be facilitated by use of version control tools [@Erdemir2020] such as git [@git20XX].

Potential users of model code and data should be able to easily identify the appropriateness of a version to their needs. Using semantic versioning [@semver20xx] conventions can signal the potential importance of an update to users of model code and data. Also informative is to clearly label the type of code versions hosted in a repository as either:

- "development" - typically the most comprehensive and up to date source code, but potentially not yet sufficiently well documented and tested for potential users to apply to purposes other than testing;

- "production" - code that has been released with a view to it being applied to its stated purposes and therefore typically required to meet defined documentation and testing standards; or

- "archive" - permanent copies of code at key milestones in its development that can aid study reproductions by making available the version of code used in a specific analysis.

Continuous integration [@CI2017] tools can help verify that each code update continues to pass a standardised battery of quality tests. Finally, using deprecation conventions that take an informative and staged approach to retiring old code and data reduces the risk of model revisions having unintended consequences on third party users.

### 2023-01-31 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
edits up to mid discussion

## Repositories <!-- Existing ref -->
We created a GitHub organisation (a collection of code repositories) where all our development code is stored and version controlled [@ready4gh2022]. We linked our GitHub organisation to an account we established at codecov [@codecov_2022] to track our code coverage. We configured the repositories in our GitHub organisation to use GitHub actions to support continuous integration. Some of the continuous integration checks we have defined assess each library's compliance with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022], to which we plan submitting future production releases.

For hosting documentation websites for the code libraries we create, we enabled GitHub Pages in each library's repository. We also developed a consolidated and versioned project documentation website [@rfwn2022], using the Hugo framework [@hugo_2023] and Docsy theme [@docsy_2023]. To host that website we established an account with Netlify [@netlify_2023] and linked that service to our GitHub organisation so that the website would automatically update whenever the source code in its GitHub repository was edited.

We also created a Zenodo community [@ready4zen2022] - a collection of permanent, uniquely identified repositories. We then linked our Zenodo community and GitHub organisation so that every time we create a new "release" in a GitHub repository, a copy of that code is automatically created on Zenodo with a DOI. Finally, to manage model datasets, we created a dedicated collection within the Harvard Dataverse installation [@ready4dv2022].

## Software <!-- Novel ref -->
We created six development version R libraries to help us author model modules, supply them with data and implement reproducible modelling analyses. The six libraries, their primary focus, the standards they support and the third-party packages they depend on are summarised in Table \@ref(tab:cpkgs).

A library called ready4 [@ready42022] defines a template module (using R's S4 type class) from which all model module data structures will inherit features and a novel syntax for implementing the algorithms attached to model modules. The ready4 library also contains tools for retrieving web based information on model modules, datasets and analysis programs and for partially automating updates to the project website.

@@

When used in conjunction with repositories, the six R libraries provide strong support for implementing 17 out of 20 framework standards. Standards relating to safe dissemination of statistical models (R8), user-interface development (R9) and deprecation conventions (U4) are better met through using existing third party R libraries. Preparing statistical models for dissemination can be accomplished with standard R data management tools like the dplyr [@dplyr2022] and purrr [@dplyr2022] libraries. User-interfaces are typically developed with the shiny [@shiny2022] library, for which a tutorial aimed at health economists is available [@SmithR2020]. The library lifeycle [@lifecycle2021] provides tools for R developers to consistently deprecate their code.

>>>>

## Repositories <!-- Existing ref -->
We created a GitHub organisation (a collection of code repositories) where all our development code is stored and version controlled [@ready4gh2022]. We configured the repositories in our GitHub organisation to use GitHub actions to support continuous integration. Some of the continuous integration checks we have defined assess each library's compliance with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022], to which we plan submitting future production releases. To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022].

We enabled GitHub Pages in each repository we used for code library development to facilitate the creation and hosting of library documentation websites. We also developed a consolidated and versioned project documentation website [@rfwn2022] using the Hugo framework [@hugo_2023] and Docsy theme [@docsy_2023]. To host that website we established an account with Netlify [@netlify_2023] and linked that service to our GitHub organisation so that the website would automatically update whenever the source code in its GitHub repository was edited.

We also created a Zenodo community [@ready4zen2022] - a collection of permanent, uniquely identified repositories. We then linked our Zenodo community and GitHub organisation so that every time we specify a version of code in one of our GitHub repositories as a "release", a copy of that code is automatically created on Zenodo with a DOI. Finally, to manage model datasets, we created a dedicated collection within the Harvard Dataverse installation [@ready4dv2022].

## Software <!-- Novel ref -->
We created six development version R libraries to help us author model modules, supply those modules with data and implement reproducible modelling analyses. The six libraries, their primary focus, the standards they support and the third-party packages they depend on are summarised in Table \@ref(tab:cpkgs).

A library called ready4 [@ready42022] defines a template module (using R's S4 class system) from which all model module data structures will inherit features and a novel syntax for attaching algorithms to those data structures. The ready4 library also contains tools for retrieving web based information on model modules, datasets and analysis programs and for partially automating updates to the project documentation website.

@@

When used in conjunction with toolkit repositories, the six R libraries provide support for implementing 17 out of 20 framework standards (Table \@ref(tab:timelygls)). Standards relating to safe dissemination of statistical models (R8), user-interface development (R9) and deprecation conventions (U4) are better met through using existing third party R libraries. Preparing statistical models for dissemination can be accomplished with standard R data management tools like the dplyr [@dplyr2022] and purrr [@dplyr2022] libraries. User-interfaces are typically developed with the shiny [@shiny2022] library, for which a tutorial aimed at health economists is available [@SmithR2020]. The library lifeycle [@lifecycle2021] provides tools for R developers to consistently deprecate their code.
